An Interpretable Machine Learning Framework for Modeling High-Resolution Spectroscopic Data\footnote{Open source code at \url{https://github.com/gully/blase}}

### 2022-09-23 paper/paper1/ms.tex
Nice discussion on missing lines!

\subsection{Identifying missing lines}\label{secMissingLines}
\edit1{So far we have not addressed the inevitable prospect that some spectral line is present in the data, but was not predicted at the time of the atmosphere simulation.  This ``missing lines'' scenario arises from our incomplete and evolving knowledge of quantum mechanics and molecular chemistry.  \texttt{ROBOSPECT} \citep{2013PASP..125.1164W} and the \emph{Cannon} \citep{2017ApJ...836....5H} have data-driven ways to identify these lines.  The \texttt{Cannon} can go further to indicate how these previously unknown lines correlate with the stellar labels, providing some physical undersanding for their cause. \emph{Blas\'e} as it currently stands has no mechanism for identifying these lines.  Here we consider two scenarios: \emph{I)} information about the lines now exists, but they were not included in the spectrum forward modeling process, or \emph{II)} the lines are truly anomalous and unknown.  For the first scenario, a Voigt profile or profiles could be instantiated on top of the line center location(s) of the previously unknown line, and it would be fitted alongside all its other fellow lines:}

@@

for a large number of stars.  One missing piece here is either to post-process the  \emph{blas\'e}-output EW catalog with chemical labels (C, N, O, Si, Ba, Eu, etc.), or incorporate that information directly into the structure of  \emph{blas\'e}.  The latter case is intriguing.    Rather than wait for PHOENIX to receive the necessary update,  \emph{blas\'e} could instantiate one or more such lines in the locations guided by the atomic line list.  This line could get initialized with some small amplitude, and would get added onto the underlying PHOENIX source as yet another freely tunable line.  In this way, all the other nearby lines would be used for assigning continuum and blending, but this free-standing lanthanum line would acquire a value properly taking into account the impact of its neighbors.

>>>>

\subsection{Identifying missing lines}\label{secMissingLines}
\edit1{So far we have not addressed the inevitable prospect that some spectral line or lines reside in the observed data spectrum, but are absent entirely from the precomputed model.  This ``missing line'' scenario arises from our incomplete and evolving knowledge of quantum mechanics and molecular chemistry.  \texttt{ROBOSPECT} \citep{2013PASP..125.1164W} and the \texttt{Cannon} \citep{2017ApJ...836....5H} have data-driven ways to identify these lines.  The \texttt{Cannon} can go further to indicate how these previously unknown lines correlate with the stellar labels, providing some physical undersanding for their cause. \emph{Blas\'e} as it currently stands has no mechanism for identifying these lines.  Here we consider that there are $N_\mathrm{missing}$ lines with two scenarios: \emph{I)} the identification, line center location, and possibly other ancillary information about the lines has become newly available,  or \emph{II)} the lines are truly anomalous and unknown.}

\edit1{For the first scenario, Voigt profiles could be instantiated at the newly-established line center locations $\lambda_{c,m}$ of the previously unknown line.  So Equation \ref{equation1} would become a product of the new and old physics:}

@@

\noindent \edit1{Scenario \emph{II}---truly unknown lines---would not have the luxury of knowing where to initialize the line center location, at least not based on precise theory.  So instead, an educated guess for $\lambda_\mathrm{c,m}$ could be made based on the data centroid.  }

\edit1{In either case, the initialization of $A_m, \sigma_m, \gamma_m$ would be a matter of taste, and they would therefore be difficult to regularize, at least in the manner that we have described so far.  For these reasons, an even better strategy may be to treat the data residuals not as lines, but in some other basis.  For example it could be possible to model the \emph{blas\'e} residual spectrum with, say, \texttt{wobble}, or with a Gaussian process.  These questions remain an open research area.}
